AttCRISPR : an spacetime interpretable model for sgRNA efficiency prediction

### 2020-10-21 Paper/main.tex
10-21 paper

\documentclass{bioinfo}
\copyrightyear{2020} \pubyear{2020}

@@

\begin{table}[!tpb]
    \processtable{The methods compared in our work and brief description of these methods\label{Tab:01}} 
    {\begin{tabular}{@{}lllp{3.5cm}l@{}}\toprule
        Method & Neural & End2end & Description\\\midrule
        CNN$^*$ & Yes & Yes & Naive convolutional neural network\\
        RNN$^*$ & Yes & Yes & Bidirectional long short-term memory neural network\\
        XGBoost$^*$ & No & Yes & Extreme Gradient Boosting regression tree\\
        MLP$^*$ & Yes & Yes & Multilayer perceptron\\
        DeepHF$^*$ & Yes & No & Bidirectional long short-term memory neural network (with hand-crafted biological features)\\
        CRISPRpred(SEQ)$^{\#}$ & No & No & A conventional machine learning pipeline\\
        SpAC & Yes & Yes & Spatial AttCRISPR\\
        TAC & Yes & Yes & Temporal AttCRISPR\\
        EnAC & Yes & Yes & Ensemble AttCRISPR (without hand-crafted biological features)\\
        StAC & Yes & No & Standard AttCRISPR\\
        \botrule
    \end{tabular}}\footnotesize\setlength{\parindent}{2em}{\emph{Note}: The method with superscript of $^*$ and $^{\#}$ is reported by \citeauthor{wang2019optimized} and \citeauthor{MuhammadRafid2020} respectively. 
We use the conclusion of corresponding paper as the performance indicators of the model directly. 
Specially, CRISPRpred(SEQ) take another set of hand-crafted sequence-based feature to improve performance.}
\end{table}

@@

\section{Results}
\begin{figure*}[!tpb]%figure4
    \centerline{\includegraphics[width=178mm]{baselinewithoutbiofeat.eps}}
    \caption{wait2write}\label{fig:06}
    \end{figure*}
We designed experiments to address the following questions:
\begin{itemize}
    \item 
    In the absence of hand-crafted biological features, whether the stacking of method in spatial domain and temporal domain can get better performance then use these method alone? 
    $\rightarrow$ Section~\ref{section:stacking}
    \item 
    How does AttCRISPR perform compared to current state-of-the-art methods, covering both conventional machine learning and deep-learning models? 
    $\rightarrow$ Section~\ref{section:comparison}
    \item 
    How can researcher understand the decisions make by AttCRISPR locally and globally, based on attention mechanisms? 
    $\rightarrow$ Section~\ref{section:interpretability}
    \vspace*{1pt}
\end{itemize}

@@

In addition, in Table~\ref{Tab:02}, the performance of other methods without using hand-crafted biological features, are also recorded. 
Regardless of the method we developed, RNN reported by \citeauthor{wang2019optimized}, which can be categorized as the method in temporal domain, 
is the most predictive with Spearman correlation coefficients of 0.856, 0.849, 0.851. 
It's obvious that the ensemble AttCRISPR is better at prediction (Figure~6\vphantom{\ref{fig:06}}a-c). 
Furthermore, the prediction ability of models could be boosted by addition of other hand-crafted biological features, which can't be obtained directly by sequence information.

Hence, the second experiment is designed to compare the performance of standard AttCRISPR (hand-crafted biological features are used to improve performance of ensemble AttCRISPR)
 and current state-of-the-art method, DeepHF (the RNN mentioned above integrated with hand-crafted biological features).
\begin{table}[!tpb]
    \processtable{wait2write\label{Tab:02}} 
    {\begin{tabular}{@{}llll@{}}\toprule
        Method & WT-SpCas9 & eSpCas9(1.1) & SpCas9-HF1\\\midrule
        XGBoost & 0.845 & 0.831 & 0.818\\
        MLP & 0.842 & 0.846 & 0.844\\
        CNN & 0.846 & 0.831 & 0.834\\
        RNN & 0.856 & 0.849 & 0.851\\
        TAC & 0.857 & 0.844 & 0.851\\
        SpAC & 0.862 & 0.854 & 0.857\\
        \textbf{EnAC} & \textbf{0.868} & \textbf{0.859} & \textbf{0.862}\\
        CRISPRpred(SEQ) & 0.838 & 0.830 & 0.821\\
        \botrule
    \end{tabular}}\footnotesize\setlength{\parindent}{2em}{\emph{Note}: wait2write}
\end{table}

@@

text

@@

\vspace*{-12pt}

@@

\bibliographystyle{natbib}
%\bibliographystyle{achemnat}
%\bibliographystyle{plainnat}
%\bibliographystyle{abbrv}
%\bibliographystyle{bioinformatics}
%
%\bibliographystyle{plain}
%
\bibliography{document}

>>>>

\documentclass{bioinfo}
\copyrightyear{2020} \pubyear{2020}
\usepackage{booktabs}

@@

\begin{table}[!tpb]
    \processtable{The methods compared in our work and brief description of these methods\label{Tab:01}} 
    {\begin{tabular}{@{}lllp{3.5cm}l@{}}\toprule
        Method & Neural & End2end & Description\\\midrule
        CNN$^*$ & Yes & Yes & Naive convolutional neural network\\
        RNN$^*$ & Yes & Yes & Bidirectional long short-term memory neural network\\
        XGBoost$^*$ & No & Yes & Extreme Gradient Boosting regression tree\\
        MLP$^*$ & Yes & Yes & Multilayer perceptron\\
        DeepHF$^*$ & Yes & No & Bidirectional long short-term memory neural network (with hand-crafted biological features)\\
        CRISPRpred(SEQ)$^{\#}$ & No & No & A conventional machine learning pipeline\\
        SpAC & Yes & Yes & Spatial AttCRISPR\\
        TAC & Yes & Yes & Temporal AttCRISPR\\
        EnAC & Yes & Yes & Ensemble AttCRISPR (without hand-crafted biological features)\\
        StAC & Yes & No & Standard AttCRISPR\\
        \botrule
    \end{tabular}}\footnotesize\setlength{\parindent}{2em}{\emph{Note}: The method with superscript of $^*$ and $^{\#}$ is reported by \citeauthor{wang2019optimized} and \citeauthor{MuhammadRafid2020} respectively. 
Specially, CRISPRpred(SEQ) take another set of hand-crafted sequence-based feature to improve performance.}
\end{table}

@@

\section{Results}
\begin{figure}[!tpb]%figure4
    \centerline{\includegraphics[width=86mm]{baselinewithoutbiofeat.eps}}
    \caption{In the absence of hand-crafted biological features, performance of different algorithms for sgRNA activity prediction. 
    (a)-(c) The performance of Temporal AttCRISPR, Spatial AttCRISPR and Ensemble AttCRISPR. 
    The half-violin plots show the mean and distribution of the Spearman correlation coefficient between predicted and measured sgRNA activity scores over all tests. 
    (e)-(g) In the absence of hand-crafted biological features, the performance of all prediction methods in these datasets as far as we know. 
    The $mean \pm s.d.$ of the Spearman correlation coefficient between predicted and measured sgRNA activity scores are shown in the bar plots.}\label{fig:06}
\end{figure}
We designed experiments to address the following questions:
\begin{itemize}
    \item 
    In the absence of hand-crafted biological features, whether the stacking of method in spatial domain and temporal domain can get better performance then use these method alone? 
    $\rightarrow$ Section~\ref{section:stacking}
    \item 
    How does AttCRISPR perform compared to current state-of-the-art methods, covering both conventional machine learning and deep-learning models? 
    $\rightarrow$ Section~\ref{section:comparison}
    \item 
    How can researcher understand the decisions make by AttCRISPR locally and globally, based on attention mechanisms? 
    $\rightarrow$ Section~\ref{section:interpretability}
    \vspace*{1pt}
\end{itemize}

@@

In addition, in Table~\ref{Tab:02}, the performance of other methods without using hand-crafted biological features, are also recorded. 
Regardless of the method we developed, RNN reported by \citeauthor{wang2019optimized}, which can be categorized as the method in temporal domain, 
is the most predictive with Spearman correlation coefficients of 0.856, 0.849, 0.851. 
It's obvious that the ensemble AttCRISPR is better at prediction (Figure~6\vphantom{\ref{fig:06}} a-c). 
Furthermore, the prediction ability of models could be boosted by addition of other hand-crafted biological features, which can't be obtained directly by sequence information.

Hence, the second experiment is designed to compare the performance of standard AttCRISPR (hand-crafted biological features are used to improve performance of ensemble AttCRISPR)
 and current state-of-the-art method, DeepHF (the RNN mentioned above integrated with hand-crafted biological features).
\begin{table}[!tpb]
    \processtable{
    Performance comparisons for different methods in the absence of hand-crafted biological features 
    (take Spearman correlation coefficient as evaluation index)\label{Tab:02}} 
    {\begin{tabular}{@{}lccc@{}}\toprule
        Method & WT-SpCas9 & eSpCas9(1.1) & SpCas9-HF1\\\midrule
        XGBoost$^*$ & 0.845 & 0.831 & 0.818\\
        MLP$^*$ & 0.842 & 0.846 & 0.844\\
        CNN$^*$ & 0.846 & 0.831 & 0.834\\
        RNN$^*$ & 0.856 & 0.849 & 0.851\\
        TAC & 0.857 & 0.844 & 0.851\\
        SpAC & 0.862 & 0.854 & 0.857\\
        EnAC & \textbf{0.868} & \textbf{0.859} & \textbf{0.862}\\
        CRISPRpred(SEQ)$^{\#}$ & 0.838 & 0.830 & 0.821\\
        \botrule
    \end{tabular}}{}
    \processtable{
    Performance comparisons for methods before and after integrating with hand-crafted biological features
(take Spearman correlation coefficient as evaluation index)\label{Tab:03}} 
    {\begin{tabular}{lcccccc}\toprule
        & \multicolumn{2}{c}{WT-SpCas9} & \multicolumn{2}{c}{eSpCas9(1.1)} & \multicolumn{2}{c}{SpCas9-HF} \\ 
        \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \noalign{\smallskip} 
        Method & Mean  & \begin{tabular}[c]{@{}l@{}}Std Dev\\ ($\times 10^{-3}$)\end{tabular} & Mean & \begin{tabular}[c]{@{}l@{}}Std Dev\\ ($\times 10^{-3}$)\end{tabular} & Mean & \begin{tabular}[c]{@{}l@{}}Std Dev\\ ($\times 10^{-3}$)\end{tabular} \\ \hline
        RNN$^*$ & 0.856 & 3.33 & 0.849 & 5.00 & 0.851 & 4.11 \\
        EnAC   & 0.868 & 2.66 & 0.859 & 4.66 & 0.862 & 3.19 \\
        DeepHF$^*$ & 0.867 & \textbf{2.37} & 0.862 & 4.24 & 0.860 & 3.21 \\
        StAC   & \textbf{0.872} & 2.55 & \textbf{0.867} & \textbf{3.71} & \textbf{0.867} & \textbf{2.65} \\
     \botrule
\end{tabular}}\footnotesize\setlength{\parindent}{2em}{\emph{Note}: 
The method with superscript of $^*$ and $^{\#}$ is reported by \citeauthor{wang2019optimized} and \citeauthor{MuhammadRafid2020} respectively. 
In the tables, we use the results reported in the relevant paper as the performance of the method directly.}
\end{table}

@@

In order to make the comparison apples to apples, 
we follow Equation~(\ref{eq:20}) to modify the ensemble method and design the control experiment using the same strategy to validate the conclusion that integrating with hand-crafted biological features can improve the predictive performance of methods. 
What's more, we compare the standard AttCRISPR and current state-of-the-art method, DeepHF, and show the results in Table~\ref{Tab:03}.

As shown in Table~\ref{Tab:03}, in the absence of hand-crafted biological features, AttCRISPR has significant advantages over DeepHF in predictability. 
Further, integration with the hand-crafted biological features can also improve the performance of AttCRISPR, and achieve Spearman correlation coefficients of 0.872, 0.867 and 0.867 for WT-SpCas9, eSpCas9(1.1) and SpCas9-HF1 respectively. Meanwhile, current state-of-the-art method, DeepHF achieve 0.867, 0.862 and 0.860 respectively. 
After integrating with biological features, the performance gap between AttCRISPR and DeepHF is shortened, while AttCRISPR still has better performance. In addition, we also compare the standard deviation of data obtained in ten tests, which are shown in Table~\ref{Tab:03}. 
It reveal that AttCRISPR is more stable than DeepHF.

\begin{figure}[!tpb]%figure4
    \centerline{\includegraphics[width=86mm]{baselinewithbiofeat.eps}}
    \caption{wait2write}\label{fig:07}
\end{figure}

@@

Text\vspace*{-12pt}

@@

\bibliographystyle{natbib}
\bibliography{document}
